---
title: "TidyTuesday: UK Baby Names"
subtitle: "Week 24 · 2026-06-16"
author: "Your Name"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    theme: flatly
execute:
  echo: true
  warning: false
  message: false
---

## Setup

```{r setup}
tt_fix_path <- function(path) {
  if (is.null(path) || is.na(path) || !nzchar(path)) {
    return(path)
  }
  if (.Platform$OS.type == "windows" && grepl("^/home/", path)) {
    path <- sub("^/home/", "//wsl.localhost/Ubuntu/home/", path)
  }
  normalizePath(path, winslash = "/", mustWork = FALSE)
}

week_dir_env <- Sys.getenv("TT_WEEK_DIR", unset = "")
project_root_env <- Sys.getenv("TT_PROJECT_ROOT", unset = "")
input_file <- ""

if (nzchar(week_dir_env) && nzchar(project_root_env)) {
  week_dir <- tt_fix_path(week_dir_env)
  project_root <- tt_fix_path(project_root_env)
} else {
  input_file <- knitr::current_input(dir = FALSE)
  if (is.null(input_file) || !nzchar(input_file)) {
    input_file <- knitr::current_input(dir = TRUE)
  }
  week_dir <- tt_fix_path(dirname(input_file))
  project_root <- tt_fix_path(file.path(week_dir, ".."))
}

source(file.path(project_root, "install_packages.R"))

library(tidyverse)

source(file.path(week_dir, "R/load_data.R"))
source(file.path(week_dir, "R/01_regional_comparison.R"))
source(file.path(week_dir, "R/02_name_uniqueness.R"))
source(file.path(week_dir, "R/03_bridgerton_trend.R"))

dir.create(file.path(week_dir, "output"), showWarnings = FALSE)

if (!file.exists(file.path(week_dir, "data/england_wales_names.csv"))) {
  download_data(data_dir = file.path(week_dir, "data"))
}

baby_names <- load_baby_names(data_dir = file.path(week_dir, "data"))
uniqueness <- compute_uniqueness(baby_names)
comparison_year <- min(2024, max(baby_names$Year, na.rm = TRUE))
```

Data from [TidyTuesday 2026-06-16](https://github.com/rfordatascience/tidytuesday/tree/main/data/2026/2026-06-16): baby names for England & Wales, Scotland, and Northern Ireland.

---

## Angle 1: Regional comparison

How do the most popular names compare across the three UK regions?

```{r regional-top}
plot_regional_top_names(baby_names, year = comparison_year)
```

```{r regional-overlap}
plot_regional_overlap(baby_names, year = comparison_year)
```

```{r regional-table}
summarise_regional_comparison(baby_names, year = comparison_year) |>
  knitr::kable(caption = paste("Most popular name by region and sex,", comparison_year))
```

**Takeaway:** Noah, Oliver, and Muhammad appear frequently in boys' top lists; girls' names show more regional variation. Names that land in multiple regional top-10 lists (e.g. Olivia, Isla) suggest shared cultural trends across the UK.

---

## Angle 2: Name uniqueness

Are boys' or girls' names more unique? We measure:

- **Distinct names per 1,000 births** — higher means more naming diversity
- **Share outside top 100** — higher means fewer parents pick ultra-common names

```{r uniqueness-time}
plot_uniqueness_over_time(uniqueness)
```

```{r uniqueness-sex}
plot_uniqueness_by_sex(uniqueness)
```

```{r uniqueness-table}
summarise_uniqueness(uniqueness) |>
  knitr::kable(digits = 1, caption = "Average uniqueness metrics in the latest available year")
```

**Takeaway:** Girls' names are typically more diverse than boys' names — more distinct names per 1,000 births and a larger share of births using names outside the top 100. This pattern holds across all three regions.

---

## Angle 3: The Bridgerton trend

National Records of Scotland reported that *Bridgerton* character names rose in 2025. Do we see the same in the data?

```{r bridgerton-trend}
plot_bridgerton_trend(baby_names)
```

```{r bridgerton-yoy}
plot_bridgerton_2024_2025(baby_names)
```

```{r bridgerton-table}
summarise_bridgerton(baby_names) |>
  knitr::kable(caption = "Bridgerton name ranks in 2024 vs 2025")
```

**Takeaway:** In Scotland, Daphne jumped from 476th to 172nd, Eloise from 124th to 91st, and Penelope from 81st to 71st between 2024 and 2025 — a clear pop-culture bump. England & Wales data currently ends in 2024, so the 2025 Bridgerton spike is most visible in Scotland and Northern Ireland.

---

## Share your work

- Hashtags: `#TidyTuesday` `#RStats` `#DataViz`
- Credit: ONS, NISRA, National Records of Scotland via [TidyTuesday](https://tidytues.day)
- Include alt text when posting charts on LinkedIn

See `LINKEDIN.md` for a draft post.
