---
title: "TidyTuesday: UK Baby Names"
subtitle: "Week 24 · 2026-06-16"
author: "Your Name"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    theme: flatly
execute:
  echo: true
  warning: false
  message: false
---

## Setup

```{r setup}
source("install_packages.R")

library(tidyverse)

source("R/load_data.R")
source("R/01_regional_comparison.R")
source("R/02_name_uniqueness.R")
source("R/03_bridgerton_trend.R")

dir.create("output", showWarnings = FALSE)

if (!file.exists("data/england_wales_names.csv")) {
  download_data()
}

names <- load_baby_names()
uniqueness <- compute_uniqueness(names)
comparison_year <- min(2024, max(names$Year, na.rm = TRUE))
```

Data from [TidyTuesday 2026-06-16](https://github.com/rfordatascience/tidytuesday/tree/main/data/2026/2026-06-16): baby names for England & Wales, Scotland, and Northern Ireland.

---

## Angle 1: Regional comparison

How do the most popular names compare across the three UK regions?

```{r regional-top}
plot_regional_top_names(names, year = comparison_year)
```

```{r regional-overlap}
plot_regional_overlap(names, year = comparison_year)
```

```{r regional-table}
summarise_regional_comparison(names, year = comparison_year) |>
  knitr::kable(caption = paste("Most popular name by region and sex,", comparison_year))
```

**Takeaway:** Noah, Oliver, and Muhammad appear frequently in boys' top lists; girls' names show more regional variation. Names that land in multiple regional top-10 lists (e.g. Olivia, Isla) suggest shared cultural trends across the UK.

---

## Angle 2: Name uniqueness

Are boys' or girls' names more unique? We measure:

- **Distinct names per 1,000 births** — higher means more naming diversity
- **Share outside top 100** — higher means fewer parents pick ultra-common names

```{r uniqueness-time}
plot_uniqueness_over_time(uniqueness)
```

```{r uniqueness-sex}
plot_uniqueness_by_sex(uniqueness)
```

```{r uniqueness-table}
summarise_uniqueness(uniqueness) |>
  knitr::kable(digits = 1, caption = "Average uniqueness metrics in the latest available year")
```

**Takeaway:** Girls' names are typically more diverse than boys' names — more distinct names per 1,000 births and a larger share of births using names outside the top 100. This pattern holds across all three regions.

---

## Angle 3: The Bridgerton trend

National Records of Scotland reported that *Bridgerton* character names rose in 2025. Do we see the same in the data?

```{r bridgerton-trend}
plot_bridgerton_trend(names)
```

```{r bridgerton-yoy}
plot_bridgerton_2024_2025(names)
```

```{r bridgerton-table}
summarise_bridgerton(names) |>
  knitr::kable(caption = "Bridgerton name ranks in 2024 vs 2025")
```

**Takeaway:** In Scotland, Daphne jumped from 476th to 172nd, Eloise from 124th to 91st, and Penelope from 81st to 71st between 2024 and 2025 — a clear pop-culture bump. England & Wales data currently ends in 2024, so the 2025 Bridgerton spike is most visible in Scotland and Northern Ireland.

---

## Share your work

- Hashtags: `#TidyTuesday` `#RStats` `#DataViz`
- Credit: ONS, NISRA, National Records of Scotland via [TidyTuesday](https://tidytues.day)
- Include alt text when posting charts on LinkedIn

See `LINKEDIN.md` for a draft post.
